---
title: "Introduction to uci: urban centrality index"
output: rmarkdown::html_vignette
bibliography: '`r system.file("REFERENCES.bib", package = "uci")`'
vignette: >
  %\VignetteIndexEntry{Introduction to uci: urban centrality index} 
  %\VignetteEngine{knitr::rmarkdown} 
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

**uci** is an `R` package to measure the centrality of cities and regions. The package implements the Urban Centrality Index (UCI) originally proposed by @pereira2013urbancentrality. UCI measures the extent to which the spatial organization of a city varies from extreme monocentric to extreme polycentric in a continuous scale, rather than considering a binary classification (*either* monocentric *or* polycentric). UCI values range from 0 to 1. Values close to 0 indicate more polycentric patterns and values close to 1 indicate a more monocentric urban form.


# Installation

For now, the package is only available from GitHub.

```{r, eval = FALSE}
remotes::install_github("ipeaGIT/uci")

```



# Demonstration on sample data

```{r, message = FALSE, warning = FALSE}
library(uci)
library(sf)
library(ggplot2)
```

## Data input

text

```{r}
data_dir <- system.file("extdata", package = "uci")

grid <- readRDS(file.path(data_dir, "grid_bho.rds"))
head(grid)
```
text

```{r, eval = requireNamespace(c("sf", "ggplot2"), quietly = TRUE), out.width = "80%", fig.width = 6, fig.height = 6}

ggplot(data = grid) +
  geom_sf(aes(fill = jobs), color = NA) +
  scale_fill_viridis_c() +
  theme_void()

```

## Calculating UCI


```{r}
df <- uci(
  sf_object = grid,
  var_name = 'jobs'
  )

head(df)
```

### Boostrap

Using boostrap to find max Venables spatial separation:

```{r}
df_bootstrap <- uci(
  sf_object = grid,
  var_name = 'jobs',
  boostrap_border = TRUE,
  showProgress = FALSE
)

head(df_bootstrap)
```

# References
